---
title: "Introduction to Data Science"
subtitle: "Homework 3"
author: 
output: html_notebook
---



# Problem 1 

Load the `tidyverse` package

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
```

In this problem you will work with vehicle fuel economy data from the US EPA (Environmental Protection Agency), ranging from 1985-2015. 

Read the data using the code below:

```{r, message=FALSE, warning=FALSE}
vehicles <- read_csv("https://raw.githubusercontent.com/reisanar/datasets/master/vehicles.csv")
```

   Variable  | Description
-------------|-----------------------------------------
`id`         | Unique EPA identifier
`make`       | Manufacturer
`model`      | Model name
`year`       | Model year
`class`      | EPA vehicle size class
`trans`      | Transmission
`drive`      | Drive train
`cyl`        | Number of cylinders
`displ`      | Engine displacement, in litres
`fuel`       | Fuel type
`hwy`        | Highway fuel economy, in mpg
`cty`        | City fuel economy, in mpg


(a) (6 points) Select the different manufacturers (makes) of the cars in this data set. Save this vector in a variable called `makes`
```{r}
makes <- unique(vehicles$make)
```

(b) (6 points) Can you determine how many different car manufacturers are represented by the data set?
```{r}
length(makes)
```

(c) (6 points) Filter the data set for vehicles manufactured in 1998.
```{r}
vehicles %>% 
  filter(year == "1998")

vehicles[as.numeric(vehicles$year) == 1998,]
```

(d) (8 points) Mutate the 1998 cars data frame to add a column `average` that has the average gas mileage (between city and highway MPG) for each car
```{r}
vehicles %>% 
  filter(year == "1998") %>% 
  mutate(average = (cty+hwy)/2)
```

(e) (8 points) Of all vehicles, what is the vehicle ID of the vehicle with the worst highway MPG?
```{r}
vehicles %>% 
  arrange(hwy) %>% 
  select(id, make, model, hwy) %>% 
  filter(hwy == min(hwy))

vehicles[which(vehicles$hwy == min(vehicles$hwy)),c('id', 'make', 'model', 'hwy')]
```

(f) (8 points) Which 2013 Acura model has the best highway MPG? Which 2015 Honda model has the best highway MPG?
```{r}
vehicles %>%
  filter(make == "Acura", year == "2013") %>%
  arrange(desc(hwy)) %>%
  select(year, make, model, hwy) %>%
  filter(hwy == max(hwy))

head(arrange(vehicles[vehicles$make == "Acura" & vehicles$year == "2013",], desc(hwy)), 1)
```

```{r}
vehicles %>%
  filter(make == "Honda", year == "2015") %>% 
  arrange(desc(hwy)) %>% 
  select(year, make, model, hwy) %>% 
  filter(hwy == max(hwy))

head(arrange(vehicles[vehicles$make == "Honda" & vehicles$year == "2015",], desc(hwy)), 1)
```



# Problem 2

Load the `nycflights13` packages. 

```{r}
library(nycflights13)
```


```{r}
head(flights)
glimpse(flights)
```

Use the `flights` dataframe to answer the following questions. 

(a) (6 points) Find all flights headed to TPA, and select only the origin, destination, and the departure time.
```{r}
flights %>% 
  filter(dest == "TPA") %>% 
  select(origin, dest, dep_time)

flights[flights$dest == "TPA", c('origin', 'dest', 'dep_time')]
```

(b) (6 points) Consider flights from JFK to MCO. What was the average, minimum, and maximum _air time_
of those flights? 
```{r}
flights %>% 
  filter(origin == "JFK" & dest == "MCO") %>% 
  summarise(mean = mean(air_time, na.rm = T), min = min(air_time, na.rm = T), max = max(air_time,na.rm = T))
  # mutate(mean = mean(air_time, na.rm = T), min = min(air_time, na.rm = T), max = max(air_time, na.rm = T)) %>%
  # select(mean, min, max) %>%
  # head(1)
 
summarise(flights, mean = mean(flights$air_time[flights$origin == "JFK" & flights$dest == "MCO"], na.rm = T), 
          min = min(flights$air_time[flights$origin == "JFK" & flights$dest == "MCO"], na.rm = T), 
          max = max(flights$air_time[flights$origin == "JFK" & flights$dest == "MCO"],na.rm = T))
```

(c) (6 points) What was the average _departure delay_ in each month? Save this as a data frame `dep_delay_by_month`. (Hint: you may have to perform a grouping operation then summarizing your data).
```{r}
flights %>% 
  group_by(month) %>% 
  summarise(avg_dep_delay = mean(dep_delay, na.rm = T)) -> dep_delay_by_month

dep_delay_by_month
``` 

(d) (8 points) Which month had the greatest average _departure delay_?
```{r}
dep_delay_by_month %>% 
  filter(avg_dep_delay == max(avg_dep_delay))
```

(e) (8 points) To which destinations were the average _arrival delays_ the highest?
```{r}
flights %>% 
  group_by(dest) %>% 
  summarise(avg_arr_delay = mean(arr_delay, na.rm = T)) %>%
  select(dest, avg_arr_delay) %>%
  arrange(desc(avg_arr_delay))
```

(f) (8 points) Which airline had the smallest average _arrival delay_?
```{r}
flights %>% 
  group_by(carrier) %>% 
  summarise(avg_arr_delay = mean(arr_delay, na.rm = T)) %>% 
  select(carrier, avg_arr_delay) %>%
  filter(avg_arr_delay == min(avg_arr_delay))
```

(g) (8 points) Find all flights that were operated by United, American, or Delta.
```{r}
airlines
```

```{r}
flights %>% 
  filter(carrier == "UA" | carrier == "AA" | carrier == "DL") %>% 
  select(carrier, year, origin, dest)
  
```

(h) (8 points) Find all the flights (if any) headed to Orlando, Tampa, or Miami with a scheduled departure time later than 8:00 PM.
```{r}
grep("Miami", airports$name)
airports[877,]
```

```{r}
flights %>% 
  filter(dest == "ORD" | dest == "TPA" | dest == "MIA") %>% 
  filter(sched_dep_time > 2000) %>% 
  select(dest, sched_dep_time)
```




